Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}

### 2017-03-07 ident_manu.tex
fixed column orders in table 2

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Wed Dec 21 10:59:02 2016 -0600, \href{https://github.com/fawda123/identifiability/commit/d3eb2370a7cc79abbb571fc9216ed07baa4c881b}{d3eb2370a7cc79abbb571fc9216ed07baa4c881b}}

@@

\subsection{Observational and structural uncertainty}

Both the relative $L1$ estimates and number of parameters that affected state variables were sensitive to variation in the initial conditions (observation effects) and structural changes in the model.  Visual comparison of sensitivity values ($L1$) for the default model setup showed that all state variables were sensitive to \textit{vmaxSi} after changing the initial conditions to April or September conditions (\cref{fig:sensalltile}).  State variables were also sensitive to \textit{Ksi} for only the September conditions. Changing the model structure from the default (simple) to the complex setup showed state variables were also sensitive to \textit{KQn} and \textit{Tref(nospA+nospZ)$_{z1}$}.  Changes in the sensitivity of individual parameters were also observed, most notably as a disproportionate increase in $L1$ for ammonium, \ac{chla}, \ac{pom}, and \ac{dom} to \textit{mA} and \textit{Tref(nospA+nospZ)$_{p1}$} using the complex model setup.  Summaries by parameter categories of the effects of initial conditions and structural changes are shown in \cref{tab:inistrchg,fig:inistrsumm}. Changes in average sensitivity using April and September initial conditions showed most variables were less sensitive to parameters in the optics, temperature, and zooplankton categories, whereas increased sensitivity was generally observed for parameters in the organic matter and phytoplankton categories.  Some exceptions were observed, such as a decrease in nitrogen sensitivity to phytoplankton parameters (\cref{tab:inistrchg}).  Effects of structural complexity were most apparent as an increase in sensitivity of state variables to phytoplankton and temperature parameters, as noted for the individual variables in \cref{fig:sensalltile}.

@@

% initial conditions evaluated 
%latex.default(inits[, -1], file = "", rowlabel = "Variables",     caption = cap_val, caption.loc = "top", label = "tab:inits",     rowname = inits[, 1])%
\begin{table}[!tbp]
\caption{Initial conditions that were varied to evaluate effects of observational uncertainty on parameter sensitivity.  April and September values are based on seasonal averages from water quality samples of the \ac{lcs}.  All units are $\mu$mol L$^{-1}$.\label{tab:inits}} 
\begin{center}
\begin{tabular}{llll}
\hline\hline
\multicolumn{1}{l}{Variables}&\multicolumn{1}{c}{Default}&\multicolumn{1}{c}{April}&\multicolumn{1}{c}{September}\tabularnewline
\hline
Dissolved Inorganic Carbon&$2.13\times 10^{3}$&$2.1\times 10^{3}$&$2.13\times 10^{3}$\tabularnewline
Ammonium&$1.09$&$1.79$&$0.35$\tabularnewline
Nitrate&$71.4$&$3.75$&$2.04$\tabularnewline
Dissolved Oygen&$172$&$194.47$&$152.57$\tabularnewline
Phosphate&$1.81$&$0.23$&$0.42$\tabularnewline
Silica&$71.4$&$5.84$&$9.77$\tabularnewline
\hline
\end{tabular}\end{center}
\end{table}

% structural conditions
%latex.default(swtch[, -1], file = "", rowlabel = "Switch type",     caption = cap_val, caption.loc = "top", label = "tab:strcs",     rowname = swtch[, 1])%
\begin{table}[!tbp]
\caption{Model switches that were used to evaluate effects of structural uncertainty on parameter sensitivity. Complete details and equations are provided as supplementary material to \citetalias{LehrterIP}.\label{tab:strcs}} 
\begin{center}
\begin{tabular}{lll}
\hline\hline
\multicolumn{1}{l}{Switch type}&\multicolumn{1}{c}{Default}&\multicolumn{1}{c}{Complex}\tabularnewline
\hline
Temperature&sigmoidal {\scriptsize \citep{Eldridge10}}&Arrenhius {\scriptsize \citep{Geider97}}\tabularnewline
Uptake&Michaelis-Menten {\scriptsize \citep{Dugdale67}}&Geider {\scriptsize \citep{Lehman75,Geider98}}\tabularnewline
Quota&Droop {\scriptsize \citep{Droop73}}&Flynn {\scriptsize \citep{Flynn03}}\tabularnewline
Chla:C&regression {\scriptsize \citep{Murrell14}}&Cloern {\scriptsize \citep{Cloern95}}\tabularnewline
Photosynthesis&photoinhibition {\scriptsize \citep{Platt80}}&nutrient dependent\tabularnewline
Specific growth rate&Leibig's minimum&nutrient dependent\tabularnewline
\hline
\end{tabular}\end{center}
\end{table}

@@

~~\scriptsize{\textit{astarOMA}}&$2$&$7.33$&$5.42$&-&$7.87$&$5.36$&-&-&$7.78$\tabularnewline
~~\scriptsize{\textit{astarOMZ}}&$2$&-&-&$1.39$&-&-&-&$4.73$&-\tabularnewline
~~\scriptsize{\textit{astarOMA}}&$3$&-&-&$3.87$&-&-&-&$10.04$&-\tabularnewline
~~\scriptsize{\textit{astarOMZ}}&$3$&$7.58$&$5.51$&-&$7.87$&$6.02$&-&-&$7.91$\tabularnewline

@@

~~\scriptsize{\textit{KG1}}&$1$&-&-&$1$&-&-&-&$1$&$1$\tabularnewline
~~\scriptsize{\textit{KG2}}&$1$&-&-&-&$1$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{KGcdom}}&$1$&-&$1$&-&-&$1$&-&-&-\tabularnewline
~~\scriptsize{\textit{KstarO2}}&$1$&-&-&-&-&-&$1$&-&-\tabularnewline

@@

~~\scriptsize{\textit{KG1}}&$2$&-&$1.12$&-&-&$1.93$&-&-&-\tabularnewline
~~\scriptsize{\textit{KG2}}&$2$&-&-&$6$&-&-&-&$13.43$&-\tabularnewline
~~\scriptsize{\textit{KGcdom}}&$2$&-&-&-&$1.39$&-&-&-&$1.47$\tabularnewline

@@

~~\scriptsize{\textit{KG2}}&$3$&-&-&-&-&$8.19$&-&-&-\tabularnewline
~~\scriptsize{\textit{KGcdom}}&$3$&-&-&-&-&-&-&$13.75$&-\tabularnewline
~~\scriptsize{\textit{KO2}}&$3$&-&-&-&$11.96$&-&-&-&$14.07$\tabularnewline

@@

~~\scriptsize{\textit{KO2}}&$4$&-&-&-&-&-&-&$14.68$&-\tabularnewline
~~\scriptsize{\textit{KstarO2}}&$4$&-&-&-&-&$10.65$&-&-&$14.08$\tabularnewline

@@

~~\scriptsize{\textit{mA}}&$1$&$1$&$1$&-&$1$&-&-&-&$1$\tabularnewline
~~\scriptsize{\textit{umax}}&$1$&-&-&$1$&-&$1$&$1$&$1$&-\tabularnewline
~~\scriptsize{\textit{ediblevector(Z1)}}&$2$&$1.13$&$1.17$&-&-&-&-&-&$1.15$\tabularnewline
~~\scriptsize{\textit{mA}}&$2$&-&-&$1.19$&-&$1.29$&-&-&-\tabularnewline
~~\scriptsize{\textit{Qc}}&$2$&-&-&-&-&-&$11.57$&-&-\tabularnewline
~~\scriptsize{\textit{umax}}&$2$&-&-&-&$1.21$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{vmaxP}}&$2$&-&-&-&-&-&-&$7.45$&-\tabularnewline
~~\scriptsize{\textit{alpha}}&$3$&-&-&$1.44$&-&$1.98$&-&-&-\tabularnewline
~~\scriptsize{\textit{ediblevector(Z1)}}&$3$&-&-&-&$2.9$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{umax}}&$3$&$2.73$&$2.11$&-&-&-&-&-&$3.26$\tabularnewline

@@

~~\scriptsize{\textit{ediblevector(Z1)}}&$4$&-&-&$2.09$&-&$4.09$&-&-&-\tabularnewline
~~\scriptsize{\textit{Qc}}&$4$&-&-&-&-&-&-&-&$4.98$\tabularnewline
~~\scriptsize{\textit{vmaxN}}&$4$&-&-&-&$4.9$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{alpha}}&$5$&-&-&-&-&-&-&-&$10.11$\tabularnewline

@@

~~\scriptsize{\textit{ZKa}}&$1$&-&-&$1$&-&$1$&$1$&$1$&-\tabularnewline
~~\scriptsize{\textit{Zumax}}&$1$&$1$&$1$&-&$1$&-&-&-&$1$\tabularnewline
~~\scriptsize{\textit{ZKa}}&$2$&-&$4.31$&-&$5.43$&-&-&-&$7.3$\tabularnewline
~~\scriptsize{\textit{ZQn}}&$2$&-&-&$3.18$&-&$6.32$&$9.76$&$8.54$&-\tabularnewline

@@

~~\scriptsize{\textit{Zumax}}&$3$&-&-&-&-&$6.93$&-&-&-\tabularnewline
~~\scriptsize{\textit{Zm}}&$4$&-&-&-&-&$11.86$&-&-&-\tabularnewline

@@

% change in sensitivity for different initial, structural conditions
%latex.default(totab_all[, -c(1, 2)], file = "", rowlabel = "State variables by parameter category",     caption = cap_val, caption.loc = "top", size = "scriptsize",     label = "tab:inistrchg", rowname = stts, rgroup = unique(cats),     n.rgroup = as.numeric(table(cats)), colheads = colnms, cgroup = c("",         "Change"), n.cgroup = c(1, 3))%
\caption{Changes in mean sensitivity of state variables by parameter categories to different initial conditions and structural components of the model.  Changes show the difference in average sensitivity from results using the default model setup. Sensitivities are based on average $L1$ (see \cref{l1}) values of the state variables to changes in parameters in each parameter category. Increases in average sensitivity are in bold. Medians and ranges of sensitivity values for the different conditions are shown in \cref{fig:inistrsumm}.\label{tab:inistrchg}} 
\begin{tabular}{llclll}
\multicolumn{1}{l}{\bfseries State variables by parameter category}&\multicolumn{1}{c}{\bfseries }&\multicolumn{1}{c}{\bfseries }&\multicolumn{3}{c}{\bfseries Change}\tabularnewline
\cline{4-6}
\multicolumn{1}{l}{}&\multicolumn{1}{c}{Default sensitivity}&\multicolumn{1}{c}{}&\multicolumn{1}{c}{Initial: April}&\multicolumn{1}{c}{Initial: September}&\multicolumn{1}{c}{Structure: complex}\tabularnewline
{\bfseries Optics}&&&&&\tabularnewline
~~Ammonium&$0.01$&&$-9.13\times 10^{-3}$&$-8.37\times 10^{-3}$&$-7.7\times 10^{-3}$\tabularnewline
~~Chlorophyll&$8.77\times 10^{-3}$&&$-6.79\times 10^{-3}$&$-6.3\times 10^{-3}$&$-7.33\times 10^{-3}$\tabularnewline
~~Dissolved Oxygen&$2.81\times 10^{-4}$&&$-2.11\times 10^{-4}$&$-1.46\times 10^{-4}$&$-1.93\times 10^{-4}$\tabularnewline
~~Irradiance&$9.1\times 10^{-3}$&&$\bm{2.56\times 10^{-3}}$&$\bm{3.1\times 10^{-3}}$&$\bm{1.57\times 10^{-3}}$\tabularnewline
~~Nitrate&$7.72\times 10^{-3}$&&$-6.46\times 10^{-3}$&$-6.2\times 10^{-3}$&$-6.8\times 10^{-3}$\tabularnewline
~~POM&$0.01$&&$-9.06\times 10^{-3}$&$-8.49\times 10^{-3}$&$-8.56\times 10^{-3}$\tabularnewline
~~DOM&$0.02$&&$-0.01$&$-0.01$&$-0.01$\tabularnewline
~~Phosphate&$3.35\times 10^{-3}$&&$-2.53\times 10^{-3}$&$-2.15\times 10^{-3}$&$-3.04\times 10^{-3}$\tabularnewline
{\bfseries Organic Matter}&&&&&\tabularnewline
~~Ammonium&$0.31$&&$\bm{0.08}$&$\bm{0.04}$&$-2.22\times 10^{-4}$\tabularnewline
~~Chlorophyll&$0.02$&&$\bm{0.12}$&$\bm{0.14}$&$-0.01$\tabularnewline
~~Dissolved Oxygen&$2.17\times 10^{-3}$&&$-6.75\times 10^{-4}$&$\bm{2.32\times 10^{-4}}$&$-1.05\times 10^{-3}$\tabularnewline
~~Irradiance&$6.63\times 10^{-3}$&&$\bm{5.34\times 10^{-3}}$&$\bm{6.79\times 10^{-3}}$&$-2.99\times 10^{-4}$\tabularnewline
~~Nitrate&$0.12$&&$\bm{0.06}$&$\bm{0.09}$&$-0.1$\tabularnewline
~~POM&$0.13$&&$\bm{0.15}$&$\bm{0.12}$&$\bm{0.04}$\tabularnewline
~~DOM&$0.14$&&$\bm{0.07}$&$\bm{0.17}$&$\bm{0.05}$\tabularnewline
~~Phosphate&$0.03$&&$\bm{0.04}$&$\bm{0.06}$&$-0.02$\tabularnewline
{\bfseries Phytoplankton}&&&&&\tabularnewline
~~Ammonium&$0.8$&&$-0.02$&$\bm{0.3}$&$\bm{65.98}$\tabularnewline
~~Chlorophyll&$1.14$&&$\bm{1.23}$&$\bm{4.38}$&$\bm{104.24}$\tabularnewline
~~Dissolved Oxygen&$0.01$&&$-4.58\times 10^{-4}$&$\bm{2.31\times 10^{-3}}$&$\bm{3.44\times 10^{-4}}$\tabularnewline
~~Irradiance&$0.02$&&$-2.04\times 10^{-3}$&$\bm{1.83\times 10^{-3}}$&$\bm{7.53\times 10^{-3}}$\tabularnewline
~~Nitrate&$0.74$&&$-0.43$&$-0.45$&$-0.64$\tabularnewline
~~POM&$0.76$&&$\bm{0.17}$&$\bm{0.46}$&$\bm{55.11}$\tabularnewline
~~DOM&$1.28$&&$\bm{0.34}$&$\bm{1.15}$&$\bm{128.93}$\tabularnewline
~~Phosphate&$0.17$&&$\bm{0.03}$&$\bm{0.05}$&$-0.14$\tabularnewline
{\bfseries Temperature}&&&&&\tabularnewline
~~Ammonium&$0.79$&&$-0.69$&$-0.69$&$\bm{272.31}$\tabularnewline
~~Chlorophyll&$0.6$&&$-0.47$&$-0.45$&$\bm{114.51}$\tabularnewline
~~Dissolved Oxygen&$0.02$&&$-0.02$&$-0.02$&$-0.01$\tabularnewline
~~Irradiance&$0.03$&&$-0.02$&$-0.02$&$\bm{0.01}$\tabularnewline
~~Nitrate&$0.3$&&$-0.15$&$-0.16$&$-0.19$\tabularnewline
~~POM&$0.86$&&$-0.68$&$-0.69$&$\bm{215.17}$\tabularnewline
~~DOM&$1.48$&&$-1.21$&$-1.2$&$\bm{900.26}$\tabularnewline
~~Phosphate&$0.16$&&$-0.11$&$-0.09$&$-0.11$\tabularnewline
{\bfseries Zooplankton}&&&&&\tabularnewline
~~Ammonium&$0.47$&&$-0.45$&$-0.44$&$-0.16$\tabularnewline
~~Chlorophyll&$0.39$&&$-0.26$&$-0.23$&$-0.2$\tabularnewline
~~Dissolved Oxygen&$0.01$&&$-0.01$&$-8.68\times 10^{-3}$&$-6.96\times 10^{-3}$\tabularnewline
~~Irradiance&$0.03$&&$-0.02$&$-0.02$&$-0.01$\tabularnewline
~~Nitrate&$1.06$&&$-1.02$&$-1.02$&$-0.98$\tabularnewline
~~POM&$0.37$&&$-0.27$&$-0.28$&$-0.18$\tabularnewline
~~DOM&$0.39$&&$-0.3$&$-0.28$&$-0.18$\tabularnewline
~~Phosphate&$0.31$&&$-0.19$&$-0.17$&$-0.27$\tabularnewline

@@

{\centering \includegraphics[width=\textwidth]{figs/sensalltile-1}

@@

\caption{Sensitivity values (L1, \cref{l1}) of all state variables to changes in parameter values using default conditions, changes in initial conditions (April, September seasonal means, \cref{tab:inits}), and changes in structural complexity (\cref{tab:strcs}). Parameters are grouped by category: optics, organic matter, phytoplankton, zooplankton, temperature, and zoplankton.  See \cref{tab:dosens} for L1 values for \ac{do} and \cref{tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens} for the other state variables.}\label{fig:sensalltile}
\end{figure}

@@

% summary of initial/structural changes on parameter sensitivity
\begin{figure}[!ht]

{\centering \includegraphics[width=\textwidth]{figs/inistrsumm-1}

\caption{Summaries of changes in parameter sensitivity values (L1, \cref{l1}) from the default model conditions.  Sensitivity was re-evaluated using initial conditions as seasonal means for April and September, and added structural complexity.  Sensitivity is summarized as the minimum, median, and maximum for each state variable separated by parameter categories. Changes in average sensitivity from the default model conditions are in \cref{tab:inistrchg}.}\label{fig:inistrsumm}
\end{figure}

>>>>

\title{Parameter sensitivity and identifiability for a biogeochemical model of hypoxia in the northern {G}ulf of {M}exico\tnoteref{mytitlenote}}
\tnotetext[mytitlenote]{Version: Mon Mar 6 11:12:53 2017 -0600, \href{https://github.com/fawda123/identifiability/commit/4a06dbf55054203e59978de8be41053a6c7b9ba1}{4a06dbf55054203e59978de8be41053a6c7b9ba1}}

@@

@@

@@

~~\scriptsize{\textit{astarOMA}}&$2$&$7.33$&$5.42$&-&$5.36$&-&$7.78$&$7.87$&-\tabularnewline
~~\scriptsize{\textit{astarOMZ}}&$2$&-&-&$1.39$&-&-&-&-&$4.73$\tabularnewline
~~\scriptsize{\textit{astarOMA}}&$3$&-&-&$3.87$&-&-&-&-&$10.04$\tabularnewline
~~\scriptsize{\textit{astarOMZ}}&$3$&$7.58$&$5.51$&-&$6.02$&-&$7.91$&$7.87$&-\tabularnewline

@@

~~\scriptsize{\textit{KG1}}&$1$&-&-&$1$&-&-&$1$&-&$1$\tabularnewline
~~\scriptsize{\textit{KG2}}&$1$&-&-&-&-&-&-&$1$&-\tabularnewline
~~\scriptsize{\textit{KGcdom}}&$1$&-&$1$&-&$1$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{KstarO2}}&$1$&-&-&-&-&$1$&-&-&-\tabularnewline

@@

~~\scriptsize{\textit{KG1}}&$2$&-&$1.12$&-&$1.93$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{KG2}}&$2$&-&-&$6$&-&-&-&-&$13.43$\tabularnewline
~~\scriptsize{\textit{KGcdom}}&$2$&-&-&-&-&-&$1.47$&$1.39$&-\tabularnewline

@@

~~\scriptsize{\textit{KG2}}&$3$&-&-&-&$8.19$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{KGcdom}}&$3$&-&-&-&-&-&-&-&$13.75$\tabularnewline
~~\scriptsize{\textit{KO2}}&$3$&-&-&-&-&-&$14.07$&$11.96$&-\tabularnewline

@@

~~\scriptsize{\textit{KO2}}&$4$&-&-&-&-&-&-&-&$14.68$\tabularnewline
~~\scriptsize{\textit{KstarO2}}&$4$&-&-&-&$10.65$&-&$14.08$&-&-\tabularnewline

@@

~~\scriptsize{\textit{mA}}&$1$&$1$&$1$&-&-&-&$1$&$1$&-\tabularnewline
~~\scriptsize{\textit{umax}}&$1$&-&-&$1$&$1$&$1$&-&-&$1$\tabularnewline
~~\scriptsize{\textit{ediblevector(Z1)}}&$2$&$1.13$&$1.17$&-&-&-&$1.15$&-&-\tabularnewline
~~\scriptsize{\textit{mA}}&$2$&-&-&$1.19$&$1.29$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{Qc}}&$2$&-&-&-&-&$11.57$&-&-&-\tabularnewline
~~\scriptsize{\textit{umax}}&$2$&-&-&-&-&-&-&$1.21$&-\tabularnewline
~~\scriptsize{\textit{vmaxP}}&$2$&-&-&-&-&-&-&-&$7.45$\tabularnewline
~~\scriptsize{\textit{alpha}}&$3$&-&-&$1.44$&$1.98$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{ediblevector(Z1)}}&$3$&-&-&-&-&-&-&$2.9$&-\tabularnewline
~~\scriptsize{\textit{umax}}&$3$&$2.73$&$2.11$&-&-&-&$3.26$&-&-\tabularnewline

@@

~~\scriptsize{\textit{ediblevector(Z1)}}&$4$&-&-&$2.09$&$4.09$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{Qc}}&$4$&-&-&-&-&-&$4.98$&-&-\tabularnewline
~~\scriptsize{\textit{vmaxN}}&$4$&-&-&-&-&-&-&$4.9$&-\tabularnewline
~~\scriptsize{\textit{alpha}}&$5$&-&-&-&-&-&$10.11$&-&-\tabularnewline

@@

~~\scriptsize{\textit{ZKa}}&$1$&-&-&$1$&$1$&$1$&-&-&$1$\tabularnewline
~~\scriptsize{\textit{Zumax}}&$1$&$1$&$1$&-&-&-&$1$&$1$&-\tabularnewline
~~\scriptsize{\textit{ZKa}}&$2$&-&$4.31$&-&-&-&$7.3$&$5.43$&-\tabularnewline
~~\scriptsize{\textit{ZQn}}&$2$&-&-&$3.18$&$6.32$&$9.76$&-&-&$8.54$\tabularnewline

@@

~~\scriptsize{\textit{Zumax}}&$3$&-&-&-&$6.93$&-&-&-&-\tabularnewline
~~\scriptsize{\textit{Zm}}&$4$&-&-&-&$11.86$&-&-&-&-\tabularnewline

@@

@@

{\centering \includegraphics[width=0.7\textwidth]{figs/sensalltile-1}

@@

\caption{Sensitivity values (L1, \cref{l1}) of all state variables to changes in a 50\% increase in parameter values. Parameters are grouped by category: optics, organic matter, phytoplankton, zooplankton, temperature, and zoplankton.  See \cref{tab:dosens} for L1 values for \ac{do} and \cref{tab:nh4sens,tab:chlsens,tab:irrsens,tab:no3sens,tab:om1sens,tab:om2sens,tab:po4sens} for the other state variables.}\label{fig:sensalltile}
\end{figure}

@@
